---
title: "Regression Models Project"
author: "Xinli"
date: "May 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

V/S (vs) and Transmission (am) are categoercial variables. Changing them to factors.  
```{r cars}
data("mtcars")
mtcars$vs <- factor(mtcars$vs)
mtcars$am <- factor(mtcars$am)
summary(mtcars)
str(mtcars)
pairs(mtcars)
```

Since we are interested in how transmissions (am) influence Miles/(US) gallon (mpg), plotting the boxplot to know the distribution of MPG for each level of transmissions. 

##Exploratory Data Analyses
```{r}
boxplot(mpg~am, data=mtcars, xlab="Transmission", ylab="Miles/(US) gallon")
```

The plot shows that manual transmission leads to more MPG. Using T test to know if there is significant difference between automatic and manual transmissions.
```{r}
t.test(mpg~am, data=mtcars)
```

The p-value is 0.01374, which is less than 0.05. So under 0.05 significant level, null hypothesis is reject. So manual transmission is better for MPG

##Fitting Models and Model Selection
Next, to know the MPG differenct between automatic and manual transmissions, fitting a model only includes variable am. 

```{r}
fit <- lm(mpg~am, data=mtcars)
par(mfrow=c(2,2))
plot(fit)
```

The residual plots show linear model contains only one variable cannot explain MPG well, so using stepwise algorithm doing model selection. AIC is used as criterion.

```{r}
stepmodel <- step(lm(mpg~., data = mtcars),direction = "both", trace = 0)
summary(stepmodel)
```
The updated model contains three variables, Weight(wt), 1/4 mile time(qsec) and Transmission(am). And it explains 84.97% of the variance.

```{r}
anova(fit,stepmodel)
```


##Residual Plot and Diagnostics
```{r}
par(mfrow=c(2,2))
plot(stepmodel)
```

- The points in the Residuals vs. Fitted plot are randomly scattered on the plot that verifies the independence condition.

- The Normal Q-Q plot consists of the points which mostly fall on the line indicating that the residuals are normally distributed.

- The Scale-Location plot consists of points scattered in a constant band pattern, indicating constant variance.

- There are some distinct points of interest (outliers or leverage points) in the left of plot, showing there are low-leverage but high-standardized residual points.

```{r}
library(car)
vif(stepmodel)
leverage <- hatvalues(stepmodel)
tail(sort(leverage),3)
influential <- dfbetas(stepmodel)
tail(sort(influential),3)
```
- VIF tells the multicollinearity of the model is not high. 

- Leverage is not that large. Using cut off value of dfbetas is 2/sqrt(n), the result suggests there are outliers.

##Summary

1. Manual transmission is better for MPG than automatic one.

2. Adjusting for Weight(wt) and 1/4 mile time(qsec), cars with manual transmission get 2.9 more mils per gallon compared to cars with automatic transmission.